Mixing Time Estimation in Reversible Markov Chains from a Single Sample Path

### 2015-10-27 arxiv-v2/appendix-empirical.tex
fixed freedman inequality

We make use of Freedman's inequality for martingales~\citep[Theorem
1.6]{Fre75}, which we state below in a simplified form.
\begin{theorem}
  \label{thm:freedman}
  Let $Y_1, Y_2, \dotsc, Y_n$ be a bounded martingale difference
  sequence with respect to the filtration $\cF = (\cF_t)_{1 \leq t
  \leq n}$ and with $|Y_t| \leq b$ for some $b>0$.
  Let $S := \sum_{t=1}^n Y_t$ and $V := \EE{Y_t^2|\cF_{t-1}}$.
  For all $t, v > 0$,
  \[
    \Pr\Brackets{
      S > \sqrt{2vt} + 2bt/3
      \,\wedge\,
      V \leq v
    }
    \leq e^{-t}
    .
  \]
\end{theorem}
\begin{corollary}
  \label{cor:freedman}
  Under the same setting as \Cref{thm:freedman}, for any $t > 0$ and
  $c > 1$,
  \[
    \Pr\Brackets{
      S > \sqrt{2cVt} + 5bt/3
    }
    \leq
    \Parens{ 1 + \ceil{\log_c(n)} }
    e^{-t}
    .
  \]
\end{corollary}
\begin{proof}
  Define $v_i := c^i b^2t/2$ for $i = 0, 1, \dotsc, \ceil{ \log_c(n)
  }$, and let $v_{-1} := -\infty$.
  Then, since $V \leq b^2n$,
  \begin{align*}
    \Pr\Brackets{
      S > \sqrt{2\max\braces{v_0,cV}t} + 2bt/3
    }
    & =
    \sum_{i=0}^{\ceil{\log_c(n)}}
    \Pr\Brackets{
      S > \sqrt{2\max\braces{v_0,cV}t} + 2bt/3
      \,\wedge\, v_{j-1} < V \leq v_j
    }
    \\
    & \leq
    \sum_{i=0}^{\ceil{\log_c(n)}}
    \Pr\Brackets{
      S > \sqrt{2\max\braces{v_0,cv_{j-1}}t} + 2bt/3
      \,\wedge\, v_{j-1} < V \leq v_j
    }
    \\
    & \leq
    \sum_{i=0}^{\ceil{\log_c(n)}}
    \Pr\Brackets{
      S > \sqrt{2v_jt} + 2bt/3
      \,\wedge\, V \leq v_j
    }
    \\
    & \leq
    \sum_{i=0}^{\ceil{\log_c(n)}}
    e^{-t}
  \end{align*}
  where the final inequality uses \Cref{thm:freedman}.
  The conclusion now follows because
  \[
    \sqrt{2cVt} + 5bt/3
    \geq \sqrt{2\max\braces{v_0,cV}t} + 2bt/3
    .
    \qedhere
  \]
\end{proof}

>>>>

We make use of a simple corollary of Freedman's inequality for
martingales~\citep[Theorem 1.6]{Fre75}.
\begin{theorem}[Freedman's inequality]
  \label{thm:freedman}
  Let $Y_1, Y_2, \dotsc, Y_n$ be a bounded martingale difference
  sequence with respect to the filtration $\cF = (\cF_i)_{1 \leq i
  \leq n}$ and with $|Y_i| \leq b$ for some $b>0$.
  Let $V := \sum_{i=1}^n \EE{Y_i^2|\cF_{i-1}}$ and $S := \max_{1 \leq
  m \leq n} \sum_{i=1}^m Y_i$.
  For all $t, v > 0$,
  \[
    \Pr\Brackets{
      S > \sqrt{2vt} + 2bt/3
      \,\wedge\,
      V \leq v
    }
    \leq e^{-t}
    .
  \]
\end{theorem}
\begin{corollary}
  \label{cor:freedman}
  Under the same setting as \Cref{thm:freedman}, for any $t > 0$ and
  $c > 1$,
  \[
    \Pr\Brackets{
      S > \sqrt{2cVt} + 5bt/3
    }
    \leq
    \Parens{ 1 + \ceil{\log_c(2n/t)} }
    e^{-t}
    .
  \]
\end{corollary}
\begin{proof}
  Define $v_i := c^i b^2t/2$ for $i = 0, 1, 2, \dotsc, \ceil{
  \log_c(2n/t) }$, and let $v_{-1} := -\infty$.
  Then, since $V \in [0,b^2n]$,
  \begin{align*}
    \Pr\Brackets{
      S > \sqrt{2\max\braces{v_0,cV}t} + 2bt/3
    }
    & =
    \sum_{i=0}^{\ceil{\log_c(2n/t)}}
    \Pr\Brackets{
      S > \sqrt{2\max\braces{v_0,cV}t} + 2bt/3
      \,\wedge\, v_{j-1} < V \leq v_j
    }
    \\
    & \leq
    \sum_{i=0}^{\ceil{\log_c(2n/t)}}
    \Pr\Brackets{
      S > \sqrt{2\max\braces{v_0,cv_{j-1}}t} + 2bt/3
      \,\wedge\, v_{j-1} < V \leq v_j
    }
    \\
    & \leq
    \sum_{i=0}^{\ceil{\log_c(2n/t)}}
    \Pr\Brackets{
      S > \sqrt{2v_jt} + 2bt/3
      \,\wedge\, V \leq v_j
    }
    \\
    & \leq
    \Parens{ 1 + \ceil{\log_c(2n/t)} }
    e^{-t}
  \end{align*}
  where the final inequality uses \Cref{thm:freedman}.
  The conclusion now follows because
  \[
    \sqrt{2cVt} + 5bt/3
    \geq \sqrt{2\max\braces{v_0,cV}t} + 2bt/3
    .
    \qedhere
  \]
\end{proof}
